Against the Planetary Interpretation of PTFO$\,$8-8695b

### 2020-04-14 paper/ms.tex
compile da paper
update paper

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
  {\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
  versus time. The beat period of 4.48 days is visible by eye.  The
  model plotted underneath the data includes 2 harmonics at the long
  period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
  period $P_{\rm s}$.
  {\it Upper middle}: Long-period signal, equal to the raw signal
  minus the short-period signal.
  {\it Lower middle}: Short-period signal, equal to the raw signal
  minus the long-period signal.
  {\it Bottom}: residual.  The data are binned from 2 to 10 minute
  cadence as a convenience for plotting and fitting.
  \label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
  Panels are as in Figure~\ref{fig:splitsignal}.
  \label{fig:splitsignalii}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f6_comp.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:corner}
	}
\end{figure*}

@@

\input{model_comparison_table.tex}
\input{model_posterior_table.tex}

We considered nine models, with the number of harmonics per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion
(Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
$N=2$ harmonics at the short 11.74$\,$hr period, and $M=2$ harmonics
at the long 11.96$\,$hr period.  All of the models have reduced
$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible,
though not perfect agreement between the data and models.

@@

In every model, the variability at the long period is a simple
sinusoid with peak-to-peak amplitude $\approx$10\%
(Figure~\ref{fig:phasefold}, top).  The variability at the short
period is always more complex.  A dip of depth $\approx$1.2\%, fit in
our model as a transit, lasts $\approx$0.75 hours.  Superposed on the
dip is a complex signal with peak-to-peak amplitude of about 4\%,
which peaks near phase 0.25, and reaches minimum brightness between
phases -0.5 and -0.25.

Outside of the primary dip, the short-period signal is relatively
smooth, at least from phases 0 to 0.5.  However the short-period
signal is asymmetric.  The flux from phases -0.5 to 0 shows what could
be a discontinuous jump, shortly after reaching minimum.  This jump
was visible in each of the nine models we considered, with different
choices for the number of harmonics.

% The periodogram of the final residual (Figure~\ref{fig:splitsignal}
% bottom row) shows a weakly significant, poorly resolved peak at
% $\approx$8 days, consistent with the visual impression in the time
% domain that there could be a weak long-period signal present.

@@

The TESS pixels are $\approx21$'' per side, and so we need to consider
whether light from neighboring stars could affect the photometry.  The
scene is shown in Figure~\ref{fig:scene}.  
The pixels used to
measure the background level are indicated with an `\texttt{X}' hatch,
and the pixels used for the final lightcurve are shown with the
`\texttt{/}' hatch.

@@

\section{Interpretation}
\label{sec:discussion}

The TESS dip does not phase up where it is supposed to...
\label{fig:o_minus_c}

@@

>>>>

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
		{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
		versus time. The beat period of 4.48 days is visible by eye.  The
		model plotted underneath the data includes 2 harmonics at the long
		period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
		period $P_{\rm s}$.
		{\it Upper middle}: Long-period signal, equal to the raw signal
		minus the short-period signal.
		{\it Lower middle}: Short-period signal, equal to the raw signal
		minus the long-period signal.
		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
		cadence as a convenience for plotting and fitting.
		\label{fig:splitsignal}
	}
\end{figure*}

%\newpage
\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
		Panels are as in Figure~\ref{fig:splitsignal}.
		\label{fig:splitsignalii}
	}
\end{figure*}
%\newpage

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f6_comp.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
		bar
		\label{fig:corner}
	}
\end{figure*}

@@

@@

\input{model_comparison_table.tex}
%\input{model_posterior_table.tex}

We considered nine models, with the number of harmonics per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion
(Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
two harmonics at the short 11.74$\,$hr period, and two harmonics
at the long 11.96$\,$hr period.  All of the models have reduced
$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible
though imperfect agreement between the data and models.

@@

In every model, the variability at the long period is a simple
sinusoid with peak-to-peak amplitude $\approx$10\%.  The variability
at the short period is always more complex.  A dip of depth
$\approx$1.2\%, fit in our model as a transit, lasts $\approx$0.75
hours.  Superposed on the dip is a complex signal with peak-to-peak
amplitude of about 4\%, which peaks near phase 0.25, and reaches
minimum brightness between phases -0.5 and -0.25.

Outside of the primary dip, the short-period signal is relatively
smooth, at least from phases 0 to 0.5.  However the short-period
signal is asymmetric.  The flux from phases -0.5 to 0 shows what could
be a discontinuous jump, shortly after reaching minimum.  This jump
was visible in each of the nine models we considered.

The periodogram of the final residual (Figure~\ref{fig:splitsignal}
bottom row) shows a weakly significant, poorly resolved peak at
$\approx$8 days, consistent with the visual impression in the time
domain that there could be a weak long-period signal present.

@@

The TESS pixels are $\approx21$'' per side, and so before making an
interptations, we need to consider whether light from neighboring
stars could have affected the photometry.  The scene is shown in
Figure~\ref{fig:scene}.  The pixels used to measure the background
level in the SPOC lightcuirve are indicated with an `\texttt{X}'
hatch, and the pixels used in the final lightcurve aperture are shown
with the `\texttt{/}' hatch.

@@

\section{Discussion}
\label{sec:discussion}

\subsection{Long period sinusoid}
The standard interpretation of sinusoidal modulations for a
pre-main-sequence M dwarf is that we are observing a rotation period.
There is no evidence to suggest that we are seeing anything different.

\subsection{Short period dip}
The dip lasts about 45 minutes, and seems to re-occur every 11.74
hours
(Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
The dip duration is roughly the same as that observed by previous
investigators \citep{van_eyken_ptf_2012,yu_tests_2015,}
The dip depth is comparable to what has been observed in red visual
bands...
%TODO: and how does it compare to IR? bigger or smaller? cite relevant
%studies.

\paragraph{Epoch}
The TESS dip does not phase up where it is supposed to...
Figure~\ref{fig:o_minus_c}

\subsection{Short period out-of-dip modulation}
If there were a giant planet transiting PTFO 8-8695, it would tidally
distort the host star, and cause ellipsoidal photometric modulations.
The amplitude of the ellipsoidal distortion for a $1\,M_{\rm Jup}$
companion would be about 1400$\,$ppm
\citep{shporer_astrophysics_2017}.  This is significantly larger than
the typical ellipsopidal modulation induced by close-in giant planets
because the host star is puffy, and still on the pre-main-sequence.
For our estimate, we assumed $R_\star = 1.39 R_\odot$, and $M_\star =
0.39 M_\odot$ \citep{van_eyken_ptf_2012}.

Our preferred model does detect a significant ellipsoidal signal,
parametrized as the ``$B_1$'' component.  The amplitude of the signal
is $0.53 \pm 0.06\%$.  Interpreted as being caused by a planet, it
would imply a minimum planet mass $M_{\rm p} \sin i$ of $3.8\,M_{\rm
Jup}$.

\section{Plausible Physical Interpretations}

@@

PTFO 8-8695 was previously thought to potentially host a hot Jupiter.
The TESS lightcurve of PTFO 8-8695 showed a number of new features,
many of which seem to disfavor the hot Jupiter interpretatation.
The TESS data showed two key pieces of evidence.
\begin{enumerate}
  \item {\it Two periodic signals.} The ``long'' signal is a 10\%
      peak-to-peak sinusoidal modulation repeating every 11.96 hours.
      The ``short'' signal is a 4\% peak-to-peak complex modulation
      repeating every 10.74 hours. It is composed of a dip, plus at
      least two harmonics. The signals beat, and therefore cannot be
      an artifact linked to data processing.
  \item {\it A dip at the wrong orbital phase.} The clearest dip in
    the ``short'' signal was consistent with recent observations by
    \citet{tanimoto_evidence_2020}, and differed from the discovery
    epoch by 5.14 hours.
\end{enumerate}

The physical mechanism responsible for all these features remains a
matter of speculation.
With that said, the TESS data support new arguments against the planetary
interpretation of PTFO 8-8695.
First, if the long signal is caused by starspot modulation, and the
short signal by a transiting planet, what causes the additional
complex modulations seen at the short, ``orbital'', period?

Similarly, if the planet truly orbits every 10.74 hours, while the
star's equator spins every 11.96 hours, the situation is clearly
Darwin unstable.

  Given the available evidence,
  PTFO$\,$8-8695 seems consistent with the
  ``transient dipping'' phenomenology observed in many young M dwarfs.
  It seems rather unlikely to be a planet.
